---
title: "Assignment 1"
author: "Nathan Taylor"
date: "September 26, 2017"
output: 
  html_document: 
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Structures

Both arrays and lists have their uses in R as ways of representing data, 
however they behave differently. An array can only include elements of a single
or homogenous nature. This requirement means that the content of an array must
be the same type (numeric, integer, character, or logical). Additionally, 
elements within array have three dimensions to them. As a result, each element
found within an array has a row value, a column value, and a shelf value. 
Arrays allow for data within R to take on three  dimensions, with each shelf
existing 'behind' or 'in front of' another. An example of an array is below 
with 3 rows, 4 columns, and two shelves:

```{r}
argh <- array(rnorm(24,50,12), dim = c(3, 4, 2))
argh
```

A list, on the other hand, has more flexibility in terms of its content because
it allows its elements to be different types (numeric, integer, character, or
logical) or even different manners of representing data in R, such as a list, 
matrix, or dataframe. The limiting factor behind a list is the fact that it 
exists only in one dimension, so its contents do not show up in table format 
with row and column values or in shelves with a shelving value. An example of 
a list can be found below:

```{r}
lat <- list(argh, "carnival", 3.14)
lat
```


# Style

```{r}
# create vector based on values, create new vector with negative values, then
# find which values in the second list are greater than zero
a <- c(5, -2, 3, -4, 1, NA)
b <- a * -1
b[b>0]

# unsure of what's going on here, but a comment belongs here describing what
# this code is doing
num <- 1
my_condition <- round(runif(1, 0, 1))
if (my_condition) {
  num <- num + 1
  print(paste("num =", num, sep=" "))
} else {
  print("false")
}

# make an integer sequence from 1 to 50
y <- 1:50

# test if a number in the value y is even
y%%2 == 0

# subset by the test above
y_even <- y%%2 == 0

# create a data frame from a matrix with column labels
x <- matrix(c(23, 34, 35, 6, 87, 39, 21, 14, 99), nrow = 3, ncol = 3)
df <- as.data.frame(x)
names(df) <- c("reading_scores",
               "math_scores",
               "writing_scores")

# create a function that finds the mean
mean <- function(x) {
  sum(x) / length(x)
}
```

# Subsetting a `data.frame`

```{r}
View(state.x77)

# NY Data, including frost and murder
state.x77["New York", ]

# NY frost & murder, respectively
state.x77["New York", c("Frost", "Murder")]

# NY Murder, by row and column number
state.x77[32, 5]

# NY Frost, by row and column number
state.x77[32, 7]
```

# Manipulating Data

```{r}
# load dplyr
library("dplyr")

# see that airquality measures wind in mph
help(airquality)

# converting mph to kph
airquality <- transmute(airquality, Ozone = Ozone, 
                                    Solar.R = Solar.R, 
                                    Wind = Wind * 1.609, 
                                    Temp = Temp, 
                                    Month = Month, 
                                    Day = Day)
airquality
```

# Stratifying

```{r}
# create database and load dplyr
cdc <- read.csv("https://www.openintro.org/stat/data/cdc.csv")
library("dplyr")

# create new variable based on differences between weight and wtdesire
cdc <- cdc %>% mutate(d_j = wtdesire - weight)

# create subgroups based on general health and smoking with means & medians
# for each
health <- group_by(cdc, genhlth, smoke100) %>%
  summarize(avg_diff = mean(d_j, na.rm = TRUE),
            med_diff = median(d_j, na.rm = TRUE))
health
```

From the results, I can conclude that people included in this data frame have 
a lower desired weight than their actual weight. This finding appears 
independent of a person's general health status or whether they have smoked
in the past 100 days or not. Additionally, the less healthy someone reports themselves to be, the more weight they wish to lose as all differences in peopls's desired weight and actual weights are negative. It appear from this cursory look that smoking does not have as significant of an effect on the difference between actual weight and desired weight as the self-reported state of general health.
